What to simulate? Inferring the right direction for mental rotation

### 2014-02-04 man/cogsci-2014/cogsci-2014.tex
Variance, not standard deviation

We define $p(\Xb\ \vert\ \Xa, \theta, h)$ to be the similarity between
$\Xb$ and a transformation of $\Xa$: $p(\Xb\ \vert\ \Xa, \theta, h):=
S(\Xb, f(\Xa, h, \theta))$.  We do not know which vertices of $\Xb$
correspond to which vertices of $\Xa$, so $S$ must marginalize over
the set of possible mappings. For brevity, let
$\mathbf{X}_m=\M\cdot{}f(\Xa, h, \theta)$ where $\M$ is a permutation
matrix. Then:
\begin{equation}
  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{2N} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma_S^2),
  \label{eq:similarity}
\end{equation}
where $2N$ is the total number of possible mappings,\footnote{It is
  $2N$ and not $N^2$ because, in polar coordinates, vertices are
  always connected to their two nearest neighbors in the $\theta$
  dimension.}and $\sigma_S=0.15$ is the standard deviation of the
similarity. Example similarity curves are shown in
Fig.~\ref{fig:model-traces}.

Additionally, we assume that the observed shapes may only be
transformed by a small amount at a time, and each transformation is
costly in terms of computational resources. Specifically, if the
current mental image is $\Xt$, then:
\begin{equation}
  \mathbf{X}_{t+1} = \left\{ \begin{array}{ll}
      f(\Xt, 0, \epsilon) &\mbox{ continue rotating by $\epsilon$,} \\
      f(\Xt, 1, 0) &\mbox{ flip,} \\
      f(\Xa, 0, 0) &\mbox{ reset, or} \\
      f(\Xa, 1, 0) &\mbox{ reset and flip,} \\
    \end{array} \right.
  \label{eq:actions}
\end{equation}
where $\epsilon\sim \left|\mathcal{N}(0, \sigma_\epsilon)\right|$ and
$\sigma_\epsilon$ is the standard deviation of the step size in
radians.

>>>>

We define $p(\Xb\ \vert\ \Xa, \theta, h)$ to be the similarity between
$\Xb$ and a transformation of $\Xa$: $p(\Xb\ \vert\ \Xa, \theta, h):=
S(\Xb, f(\Xa, h, \theta))$.  We do not know which vertices of $\Xb$
correspond to which vertices of $\Xa$, so $S$ must marginalize over
the set of possible mappings. For brevity, let
$\mathbf{X}_m=\M\cdot{}f(\Xa, h, \theta)$ where $\M$ is a permutation
matrix. Then:
\begin{equation}
  S(\Xb, f(\Xa, h, \theta)):=\frac{1}{2N} \sum_{\M} \prod_{n=1}^N \mathcal{N}(\mathbf{x}_{bn}\ \vert \ \mathbf{x}_{mn}, \I\sigma_S^2),
  \label{eq:similarity}
\end{equation}
where $2N$ is the total number of possible mappings,\footnote{It is
  $2N$ and not $N^2$ because, in polar coordinates, vertices are
  always connected to their two nearest neighbors in the $\theta$
  dimension.}and $\sigma_S^2=0.15$ is the variance of the
similarity. Example similarity curves are shown in
Fig.~\ref{fig:model-traces}.

Additionally, we assume that the observed shapes may only be
transformed by a small amount at a time, and each transformation is
costly in terms of computational resources. Specifically, if the
current mental image is $\Xt$, then:
\begin{equation}
  \mathbf{X}_{t+1} = \left\{ \begin{array}{ll}
      f(\Xt, 0, \epsilon) &\mbox{ continue rotating by $\epsilon$,} \\
      f(\Xt, 1, 0) &\mbox{ flip,} \\
      f(\Xa, 0, 0) &\mbox{ reset, or} \\
      f(\Xa, 1, 0) &\mbox{ reset and flip,} \\
    \end{array} \right.
  \label{eq:actions}
\end{equation}
where $\epsilon\sim \left|\mathcal{N}(0, \sigma_\epsilon^2)\right|$
and $\sigma_\epsilon^2$ is the variance of the step size in radians.
